Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-08-23 chatbot_classification/aaai_2021/paper.ver_03.tex
Version 3. AAAI format.

%% LyX 2.3.1-1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass{article}
\usepackage[latin9]{inputenc}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subfigure}

\makeatletter

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
% Template for GlobalSIP-2018 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\usepackage{spconf}

@@

\makeatother

\begin{document}
\title{     
  %SFC: Few-Shot Text Classification via Similarity Fused with Classification System
  Joint training of classification model and similarity model for low-resource
  text classification in chatbot
}
\name{Jingwen Huang}   
\address{     
  \{hanmei613\}@gmail.com   
}
\maketitle

\begin{abstract}
  Building  conversational  chatbot  system  has  become  a  popular solution to customer  services  under various business scenarios. A
  conversational  chatbot needs to detect user's intent given a few words, which
  essentially  equals  to  short-text  classification  problem  in  the field of Natural  Language  Processing.  Moreover,  each  time  for  a  new service, the task-specific  chatbot  system  often needs to perform well in few-shot setups due  to lack of domain-specific data, which is quite a challenge for single-model system such as text classification model system or sentence-pair semantic similarity model system under such a low-resource condition. Therefore, in this paper, we propose
  SFC,  a 2-stage joint system with multi-task training technique for both text classification task and sentence-pair semantic similarity task to
  overcome  this challenge. Furthermore, we additionally propose an improved version of SFC to allow text classification model and sentence-pair model be combined into a joint model organized in hierarchical structure. We also
  conduct extensive experiments on 4 public and 1 private datasets in few-shot setup (i.e.,
  with only 5 to 20 training data per class). The experimental results show that
  our  system can steadily outperform several competitive single-model baselines by 2 percent
  in average accuracy.
\end{abstract}
%\keywords{}

\section{Introduction}
\label{sec:intro}
% latest version start
Task-specific conversational chatbot\cite{wen2016network} are designed to share the working pressure of human customer service agents who are responsible for solving customers' questions or queries about certain products or services. Regardless of whether the conversation is single-turn or multi-turn, the essential technical solution behind the chatbot is intent classification model which can identify the correct intent behind user's input text and find the corresponding answer.

@@

\begin{figure}[htbp]
\centering 
\subfigure[classification model]{
\label{Fig.1.sub.1}
\includegraphics[scale=0.7]{picture/picture1.pdf}}
\subfigure[similarity model]{
\label{Fig.1.sub.2}
\includegraphics[scale=0.55]{picture/picture2.pdf}}
\caption{Two popular structures of building task-specific conversational chatbot}
\label{Fig.1.main}
\end{figure}

@@

\begin{figure}[htbp]
\begin{centering}
\includegraphics[scale=0.48]{picture/picture3}
\par\end{centering}
\caption{\textbf{Network Structure of 2-stage SFC:} a joint system of multi-task for both text classification task and sentence-pair semantic similarity task.}
\label{fig:ver.1.3}
\end{figure}

@@

\begin{figure*}[t]
\begin{centering}
\includegraphics[scale=0.72]{picture/picture4} 
\par\end{centering}
\caption{\textbf{Network Structure of hierarchical SFC:} a joint model of sentence-pair model and text classification model organized in hierarchical structure by completing different tasks at different network layers.}
\label{fig:framework}
\end{figure*}

\section{Methodology}
In this section, we will describe a system of similarity model fused with classification model, named as SFC, in detail. We come up with 2 different approaches for implementation of SFC system, in which hierarchical SFC is a further improvement of 2-stage SFC for better performance. 
\subsection{2-stage SFC with multi-task learning}
The detailed network structure of 2-stage SFC is shown in Fig. \ref{fig:ver.1.3}. It is composed with 2 separate stages in charge of different jobs.
In stage 1, we can use any auxiliary model or tool to provide top-K most related class labels for sampling sentence pairs of both positive samples and negative samples for stage 2. Suppose we have a training data set $\mathcal{D}=\{x_{i},y_{i}\}_{i=1}^{N}$ of $N$ data points, in which $x_{i}$ is the user input query and $y_{i}$ is a single class label from a class label set of $C$ classes in total. If we want to generate sentence pair data set $\mathcal{D'}=\{[(x,x')_{j},l_{j}]\}_{j=1}^{M}$, where $(x, x')_{j}$ is sentence pair grouped by 2 user input queries and $l_{j}\in\{0, 1\}$ is similarity class label, see Equ. \ref{eq:similarity label}, 
\begin{align}
l_{j}=\begin{cases} \label{eq:similarity label}
0 & y_{x}\neq y_{x'}\\
1 & y_{x}=y_{x'}
\end{cases}
\end{align}
directly from original data set $\mathcal{D}$ without any sampling strategy, then $\mathcal{D'}$ will have $M=N^2$ data points of sentence pairs, which might be too time-consuming for the training process of stage 2 since we often have hundreds of class labels under task-specific chatbot scenario. Although we may only have a few data points for a single class label, $M$ is still quite a big number.

@@

\subsubsection{stage 1: classification model for providing top-K candidate class labels}
As for stage 1, our SFC system can adapt to any auxiliary model or even searching tool based on Term Frequency-Inverse
Document Frequency (TF-IDF) retrieval. In this paper, however, we choose to fine-tune RoBERTa on our task-specific data set $\mathcal{D}$ by adding a softmax classifier on top of it, the same as the structure shown in Fig. \ref{Fig.1.sub.1}, for better overall performance. According to our experimental result, RoBERTa can guarantee that approximately 95\% ground truth label be included in top-5 candidate class labels with highest probability after being fine-tuned on task-specific data set with experimental result measured by F-score at around 0.75. So we use RoBERTa as encoder of classification model in our setting at stage 1 to ensure low limitation for the performance of stage 2.

@@

\subsubsection{stage 2: sentence-pair model for multi-task learning}
As for stage 2, we also choose RoBERTa as the encoder for sentence pairs, as shown in Fig. \ref{fig:ver.1.3}. Before sending sentence pairs to Roberta, we pack sentence pairs into a single sequence and separate them with a special token [SEP]. Roberta will also add a learned embedding to every token indicating whether it belongs to the first sentence or the second one. Then, we utilize the cross-attention mechanism inside RoBERTa, and then extract semantic relationship between 2 sentences from the output softmax layer, which is placed on top of the last hidden representation of the [CLS] token.

@@

\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Dataset & Domain & \#Class & \#Samples & \#Samples/Class & Experimental Settings\tabularnewline
\hline 
ITG & FAQ chatbot & 228 & 3938 & 17 & 3-fold\tabularnewline
\hline 
Amazon-670k & Product review & 250 & 1942 & 8 & class sampling\tabularnewline
\hline 
HWU64 & Intent detection & 64 & [320, 640, 960, 1280] & [5, 10, 15, 20] & data sampling\tabularnewline
\hline 
CLINC150 & Intent detection & 150 & [750, 1500, 2250] & [5, 10, 15] & data sampling\tabularnewline
\hline 
BANKING77 & Intent detection & 77 & [385, 770, 1155, 1540] & [5, 10, 15, 20] & data sampling\tabularnewline
\hline 
\end{tabular}
\par\end{centering}
\caption{Key statistics for experimental datasets and few shot settings}

@@

\par\end{centering}
\caption{F-scores on five task-specific dataset for text classification in chatbot under low resource. For ITG, we keep the full dataset. For Amazon-670k, we randomly sampled 250 classes with training sample numbers within 5-15 samples per class. For CLINC150, BANKING77, HWU64, we set up various few-shot settings (5/10/15/20 samples per class) while keeping the test set to be fixed. The highest scores for each data setting are marked in bold.}

@@

\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
\multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
\multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{ITG}}& 2-stage SFC & $0.8034$ & $0.8124$ & $0.8008$ & $0.7967$ & $0.7934$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7882$ & $0.8114$ & $0.8010$ & $0.7972$ & $0.7826$\tabularnewline
\hline
\multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& 2-stage SFC & $0.7278$ & $0.7364$ & $0.7366$ & $0.7328$ & $0.7204$\tabularnewline
\cline{2-7}
\multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7334$ & $0.7445$ & $0.7516$ & $0.7344$ & $0.7373$\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\caption{We show the performance of SFC from different settings of hyperparameters for $K$, the candidate class number, and $P$, the sample number for each class.}

@@

% latest version end

%\section{Attention based neural network}

%\section{Conclusion}

\bibliographystyle{IEEEbib}
\bibliography{refs}

>>>>

\def\year{2021}\relax
%File: formatting-instructions-latex-2021.tex
%release 2021.1
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai21}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
%\nocopyright
%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
\pdfinfo{
  /Title (AAAI Press Formatting Instructions for Authors Using LaTeX -- A Guide)
  /Author (AAAI Press Staff, Pater Patel Schneider, Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen, Francisco Cruz, Marc Pujol-Gonzalez)
  /TemplateVersion (2021.1)
}

\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subfigure}
\usepackage{spconf}

\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.

@@

@@

\title{     
  %SFC: Few-Shot Text Classification via Similarity Fused with Classification System
  Joint training of classification model and similarity model for low-resource
  text classification in chatbot
}

\author{
  %Authors
  % All authors must be in the same font size and format.
  Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
  AAAI Style Contributions by Pater Patel Schneider,
  Sunil Issar,  \\
  J. Scott Penberthy,
  George Ferguson,
  Hans Guesgen,
  Francisco Cruz,
  Marc Pujol-Gonzalez
  \\
}
\affiliations{
  %Afiliations

  \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
  %If you have multiple authors and multiple affiliations
  % use superscripts in text and roman font to identify them.
  %For example,

  % Sunil Issar, \textsuperscript{\rm 2}
  % J. Scott Penberthy, \textsuperscript{\rm 3}
  % George Ferguson,\textsuperscript{\rm 4}
  % Hans Guesgen, \textsuperscript{\rm 5}.
  % Note that the comma should be placed BEFORE the superscript for optimum readability

  2275 East Bayshore Road, Suite 160\\
  Palo Alto, California 94303\\
  % email address must be in roman text type, not monospace or sans serif
  publications21@aaai.org

  % See more examples next
}

\begin{document}
  \maketitle

  \begin{abstract}
    Building  conversational  chatbot  system  has  become a popular solution to
    customer services under various business scenarios. A conversational chatbot
    needs to detect user's intent given a few words, which essentially equals to
    short-text   classification   problem  in  the  field  of  Natural  Language
    Processing. Moreover, each time for a new service, the task-specific chatbot
    system  often  needs  to  perform  well  in  few-shot  setups due to lack of
    domain-specific  data,  which  is  quite a challenge for single-model system
    such   as   text  classification  model  system  or  sentence-pair  semantic
    similarity  model  system under such a low-resource condition. Therefore, in
    this  paper, we propose SFC, a 2-stage joint system with multi-task training
    technique  for  both  text  classification  task  and sentence-pair semantic
    similarity  task  to  overcome  this challenge. Furthermore, we additionally
    propose  an  improved  version of SFC to allow text classification model and
    sentence-pair model be combined into a joint model organized in hierarchical
    structure.  We  also conduct extensive experiments on 4 public and 1 private
    datasets  in  few-shot  setup  (i.e.,  with  only  5 to 20 training data per
    class).   The  experimental  results  show  that  our  system  can  steadily
    outperform  several  competitive  single-model  baselines  by  2  percent in
    average accuracy.
  \end{abstract}

  \section{Introduction}
  \label{sec:intro}
  Task-specific  conversational  chatbot\cite{wen2016network}  are  designed  to
  share   the  working  pressure  of  human  customer  service  agents  who  are
  responsible for solving customers' questions or queries about certain products
  or  services.  Regardless  of  whether  the  conversation  is  single-turn  or
  multi-turn,  the  essential  technical  solution  behind the chatbot is intent
  classification model which can identify the correct intent behind user's input
  text and find the corresponding answer.

@@

  \begin{figure}[htbp]
    \centering 
    \subfigure[classification model]{
      \label{Fig.1.sub.1}
      \includegraphics[scale=0.7]{picture/picture1.pdf}
    }
    \subfigure[similarity model]{
      \label{Fig.1.sub.2}
      \includegraphics[scale=0.55]{picture/picture2.pdf}
    }
    \caption{Two popular structures of building task-specific conversational chatbot}
    \label{Fig.1.main}
  \end{figure}

@@

  \begin{figure}[htbp]
    \begin{centering}
      \includegraphics[scale=0.48]{picture/picture3}
      \par
    \end{centering}
    \caption{\textbf{Network Structure of 2-stage SFC:} a joint system of multi-task for both text classification task and sentence-pair semantic similarity task.}
    \label{fig:ver.1.3}
  \end{figure}

@@

  \begin{figure*}[t]
    \begin{centering}
      \includegraphics[scale=0.72]{picture/picture4} 
      \par
    \end{centering}
    \caption{
      \textbf{Network   Structure   of  hierarchical  SFC:}  a  joint  model  of
      sentence-pair   model   and   text   classification   model  organized  in
      hierarchical  structure by completing different tasks at different network
      layers.
    }
    \label{fig:framework}
  \end{figure*}

  \section{Methodology}
  In  this  section,  we  will  describe a system of similarity model fused with
  classification  model,  named  as  SFC, in detail. We come up with 2 different
  approaches  for  implementation  of SFC system, in which hierarchical SFC is a
  further improvement of 2-stage SFC for better performance.

  \subsection{2-stage SFC with multi-task learning}

  The   detailed   network   structure   of   2-stage   SFC  is  shown  in  Fig.
  \ref{fig:ver.1.3}.  It  is  composed  with  2  separate  stages  in  charge of
  different  jobs. In stage 1, we can use any auxiliary model or tool to provide
  top-K  most  related class labels for sampling sentence pairs of both positive
  samples  and negative samples for stage 2. Suppose we have a training data set
  $\mathcal{D}=\{x_{i},y_{i}\}_{i=1}^{N}$  of  $N$ data points, in which $x_{i}$
  is the user input query and $y_{i}$ is a single class label from a class label
  set  of  $C$  classes  in total. If we want to generate sentence pair data set
  $\mathcal{D'}=\{[(x,x')_{j},l_{j}]\}_{j=1}^{M}$,   where   $(x,   x')_{j}$  is
  sentence  pair  grouped  by  2  user  input  queries and $l_{j}\in\{0, 1\}$ is
  similarity class label, see Equ. \ref{eq:similarity label},
  \begin{align}
    l_{j}=
    \begin{cases} \label{eq:similarity label}
      0 & y_{x}\neq y_{x'}\\
      1 & y_{x}=y_{x'}
    \end{cases}
  \end{align}
  directly  from  original data set $\mathcal{D}$ without any sampling strategy,
  then  $\mathcal{D'}$  will  have  $M=N^2$ data points of sentence pairs, which
  might be too time-consuming for the training process of stage 2 since we often
  have  hundreds  of class labels under task-specific chatbot scenario. Although
  we  may  only  have  a  few data points for a single class label, $M$ is still
  quite a big number.

@@

  \subsubsection*{
    stage 1: classification model for providing top-K candidate class labels
  }

  As for  stage  1,  our  SFC  system  can adapt to any auxiliary model or even
  searching  tool  based  on  Term Frequency-Inverse Document Frequency (TF-IDF)
  retrieval.  In  this  paper,  however,  we  choose to fine-tune RoBERTa on our
  task-specific  data set $\mathcal{D}$ by adding a softmax classifier on top of
  it,  the  same  as  the  structure shown in Fig. \ref{Fig.1.sub.1}, for better
  overall  performance.  According  to  our  experimental  result,  RoBERTa  can
  guarantee  that  approximately  95\%  ground  truth label be included in top-5
  candidate  class  labels  with  highest  probability after being fine-tuned on
  task-specific  data set with experimental result measured by F-score at around
  0.75.  So  we use RoBERTa as encoder of classification model in our setting at
  stage 1 to ensure low limitation for the performance of stage 2.

@@

  \subsubsection*{
    stage 2: sentence-pair model for multi-task learning
  }

  As for  stage 2, we also choose RoBERTa as the encoder for sentence pairs, as
  shown  in Fig. \ref{fig:ver.1.3}. Before sending sentence pairs to Roberta, we
  pack  sentence  pairs  into a single sequence and separate them with a special
  token  [SEP].  Roberta  will  also  add  a  learned  embedding  to every token
  indicating  whether  it belongs to the first sentence or the second one. Then,
  we  utilize  the  cross-attention  mechanism  inside RoBERTa, and then extract
  semantic relationship between 2 sentences from the output softmax layer, which
  is placed on top of the last hidden representation of the [CLS] token.

@@

      \begin{tabular}{|c|c|c|c|c|c|}
        \hline 
        Dataset & Domain & \#Class & \#Samples & \#Samples/Class & Experimental Settings\tabularnewline
        \hline 
        ITG & FAQ chatbot & 228 & 3938 & 17 & 3-fold\tabularnewline
        \hline 
        Amazon-670k & Product review & 250 & 1942 & 8 & class sampling\tabularnewline
        \hline 
        HWU64 & Intent detection & 64 & [320, 640, 960, 1280] & [5, 10, 15, 20] & data sampling\tabularnewline
        \hline 
        CLINC150 & Intent detection & 150 & [750, 1500, 2250] & [5, 10, 15] & data sampling\tabularnewline
        \hline 
        BANKING77 & Intent detection & 77 & [385, 770, 1155, 1540] & [5, 10, 15, 20] & data sampling\tabularnewline
        \hline 
      \end{tabular}
      \par
    \end{centering}
    \caption{Key statistics for experimental datasets and few shot settings}

@@

      \par
    \end{centering}
    \caption{
      F-scores  on five task-specific dataset for text classification in chatbot
      under low resource. For ITG, we keep the full dataset. For Amazon-670k, we
      randomly  sampled  250  classes  with  training sample numbers within 5-15
      samples  per  class.  For  CLINC150,  BANKING77,  HWU64, we set up various
      few-shot  settings  (5/10/15/20  samples per class) while keeping the test
      set  to  be  fixed. The highest scores for each data setting are marked in
      bold.

@@

      \begin{tabular}{|c|c|c|c|c|c|c|}
        \hline 
        \multicolumn{1}{|c|}{\multirow{2}*{Dataset}}&\multicolumn{1}{c|}{\multirow{2}*{Model}}& $K=3$ & $K=5$ & $K=10$ & $K=15$ & $K=20$\tabularnewline
        \multicolumn{1}{|c|}{} & \multicolumn{1}{c|}{} & $P=20$ & $P=10$ & $P=5$ & $P=4$ & $P=3$\tabularnewline
        \hline
        \multicolumn{1}{|c|}{\multirow{2}*{ITG}}& 2-stage SFC & $0.8034$ & $0.8124$ & $0.8008$ & $0.7967$ & $0.7934$\tabularnewline
        \cline{2-7}
        \multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7882$ & $0.8114$ & $0.8010$ & $0.7972$ & $0.7826$\tabularnewline
        \hline
        \multicolumn{1}{|c|}{\multirow{2}*{Amazon-670k}}& 2-stage SFC & $0.7278$ & $0.7364$ & $0.7366$ & $0.7328$ & $0.7204$\tabularnewline
        \cline{2-7}
        \multicolumn{1}{|c|}{} & Hierarchical SFC & $0.7334$ & $0.7445$ & $0.7516$ & $0.7344$ & $0.7373$\tabularnewline
        \hline
      \end{tabular}
      \par
    \end{centering}
    \caption{
      We  show the performance of SFC from different settings of hyperparameters
      for  $K$,  the candidate class number, and $P$, the sample number for each
      class.
    }

@@

  \bibliographystyle{aaai21}
  \bibliography{refs}
